Sets

### 2023-10-01 rec06/paper.tex
subsets

Set theory made its debut in Cantor's 1874 paper: ``On a Property of the Collection of All Real Algebraic Numbers.''
The Cantorian set theory contained several contradictions such as Russell's paradox, but mathematicians did not abandon set theory entirely.
The work of Zermelo, Fraenkel, and Skolem resulted in a well-known axiomatizations of set theory.

@@

The atomic objects in set theory are sets.
A set can be thought of as a collection of dots.
These dots have no internal structures, so the only sensible question that one can ask is ``is $x$ a member of $y$?''

The language of set theory contains a binary predicate symbol $\in$.
This is known as the \emph{membership relation}.
$x \in y$ means $x$ is an element of $y$.

@@

@@

Given a set $w$, there is a \emph{subcollection} of $w$ whose elements satisfy certain property $\varphi$.
\[
  \{ x \in w \mid \varphi(x) \}
\]
We declare that such subcollections are sets.
For example, given the set of all natural numbers $\dN$,
\[
  \{ x \in \dN \mid \mathrm{even}(x) \}
\]
is a subcollection of all even natural numbers.

>>>>

Set theory made its debut in Cantor's 1874 paper: ``On a Property of the Collection of All Real Algebraic Numbers.''
Later, mathematicians such as Russell realized that the Cantorian set theory contained several contradictions, but mathematicians did not abandon set theory.
The work of Zermelo, Fraenkel, and Skolem resulted in a well-known axiomatization of set theory.

@@

Intuitively, a set is a collection of elements.
These elements have no internal structures, so you can think of a set as a collection of dots, and the only question that one can ask is ``is $x$ a member of $y$?''
In other words, the language of set theory contains a binary predicate symbol $\in$.
This is known as the \emph{membership relation}.
$x \in y$ means $x$ is an element of $y$.

\begin{eg}
  \begin{enumerate}
  \item[]
  \item The empty set: $\varnothing$.
  \item The set containing the empty set: $\{\varnothing\}$.
  \item A set containing 3 elements: $\{ a,b,c \}$.
  \item The set of all natural numbers: $\dN = \{ 0,1,2,\ldots \}$.
  \item The set of all integers: $\dZ = \{ \ldots,-2,-1,0,1,2,\ldots \}$.
  \item The set containing $\dN$ and $\dZ$: $\{ \dN,\dZ \}$.
  \end{enumerate}
\end{eg}

\begin{eg}
  \begin{enumerate}
  \item[]
  \item Nothing is in the empty set: $x \notin \varnothing$.
  \item The set containing the empty set has an element: $\varnothing \in \{\varnothing\}$.
  \item $a \in \{ a, b, c \}$, $b \in \{ a, b, c \}$, $c \in \{ a, b, c \}$.
  \item $\dN \in \{ \dN, \dZ \}$, $\dZ \in \{ \dN, \dZ \}$, but $0 \notin \{ \dN,\dZ \}$.
  \end{enumerate}
\end{eg}

@@

\begin{eg}
  The empty set is a subset of any set.
\end{eg}
\begin{proof}
  Let $y$ be any set.
  By definition, $\varnothing \subseteq y \cng \forall z.(z \in \varnothing \imp z \in y)$.
  Let $z$ be given.
  Assume that $z \in \varnothing$.
  This allows us to derive a contradiction since $z \notin \varnothing$.\footnote{This is not a proof by contradiction, but let's ignore the details for now.}
\end{proof}

@@

Given a set $w$, there is a \emph{subcollection} of $w$ whose elements satisfy certain property $\varphi$.
\[
  \{ x \in w \mid \varphi(x) \}
\]
We declare that such subcollections are sets.
For example, given the set of all natural numbers $\dN$,
\[
  \{ x \in \dN \mid \mathrm{even}(x) \}
\]
is the subset of all even natural numbers.
